
\section{The LEDA graph input/output format}\label{leda-format}

The following passage describes the format of the output produced by the
function {\tt graph::write(ostream\& out)}. The output consists of several lines
which are separated by endl. Comment-lines have a {\tt \#} character
in the first column and are ignored. The output can be partitioned in three
sections:

\paragraph{Header Section}\ \\
The first line always contains the string {\tt LEDA.GRAPH}. If the graph
type is not parameterized, i.e. graph or ugraph, the following two
lines both contain the string {\tt void}. In case the graph is parameterized,
i.e. {\tt GRAPH} or {\tt UGRAPH}, these lines contain a description
of the node type and the edge type, which is obtained by calling the
macro {\tt LEDA\_TYPE\_NAME}.The fourth line specifies if the graph is either 
directed (-1) or undirected (-2).

\paragraph{Nodes Section}\ \\
The first line contains n, the number of nodes in the graph. The nodes
are ordered and numbered according to their position in the node list
of the graph. Each of the following n lines contains the information
which is associated with the respective node of the graph. When the
information of a node (or an edge) is sent to an output stream, it is
always enclosed by the strings {\tt |\{} and {\tt \}|}. If the graph is not
parameterized, then the string between these parantheses is empty, so
that all the n lines contain the string {\tt |\{\}|}.

\paragraph{Edges Section}\ \\
The first line contains m, the number of edges in the graph. The edges
of the graph are ordered by two criteria: first according to the number
of their source node and second according to their position in the
adjacency list of the source node. Each of the next m lines contains
the description of an edge which consists of four space-separated
parts:
\begin{description}
\item{(a)} the number of the source node
\item{(b)} the number of the target node
\item{(c)} the number of the reversal edge or 0, if no such edge is set
\item{(d)} the information associated with the edge (cf. nodes section)
\end{description}

{\bf Note}:
For the data type {\tt planar\_map} the order of the edges is important,
because the ordering of the edges in the adjacency list of a node
corresponds to the counter-clockwise ordering of these edges around the node
in the planar embedding. And the information about reversal edges is also
vital for this data type.



